---
title: "R Data Types and Functions Demo"
author: "Scott Schwartz"
date: "Week 2 Demo"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE)
```

The purpose of this demo is to introduce R data types and functions:

1. [ ] ***Variables*** like `my_vec_var` and ***assignments*** with `<-` (and `->` and `=`)
2. [ ] ***Vectors*** and the `c()` ***concatenate/combine*** function: `c(1,"2")`
3. [ ] ***Data types*** (`1`,`1.1`, `TRUE`, `"1"`) and `class()` and `typeof()` functions
<!-- https://stackoverflow.com/questions/62413619/what-is-the-difference-between-typeof-and-class-functions-and-when-are-they -->
4. [ ] ***Coercion*** and `as.numeric()`, `as.logical()`, `character()` functions
5. [ ] ***Ordinal*** with `as.integer()` versus ***nominal*** with `as.factor()`
6. [ ] ***Tibbles*** `read_csv()` versus ***Data Frames*** `read.csv()`
7. [ ] **Columns** are **variables**  (but not "R variables"!): specific things to be observed and recorded 
8. [ ] **Rows** are **observational units** (on which variables are observed and recorded)
9. [ ] A **Row** is the **observation** 
       
       - The value of a specific variable for a specific row is one piece of the  
         collection of information on an entire row which comprise an **observation** 
       
10. [ ] Beyond `glimpse()` and `head()` with `names()`, `[]`, `[[]]`, and `$` "functions"
11. [ ] The `mean`, `median`, [`mode`](https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode), `range`, `IQR`, `var`, and `sd` functions (and `%>%`)
12. [] ***Boolean Vector Mask*** selection


```{r}
```


```{r}
library(tidyverse)
coffee_ratings <- read_csv("coffee_ratings.csv")
```

# Extra Material 

## Topic 1: ***ggplot2 syntax review***

### (a) What are the first two lines of this code doing?

```{r, eval=TRUE, message=TRUE}
library(tidyverse)
coffee_ratings <- read_csv("coffee_ratings.csv")
coffee_ratings %>%
  ggplot(aes(x=total_cup_points)) +
  geom_histogram(bins=30, color="black", fill="light blue")
```

### (b) Can you read the finel three lines of the code above? What do you roughly think it's doing?

### (c) What parts of the code are parameters and what are parts are arguments?

### (d) What does the `%>%` seem to be doing in the code above?

### (e) What does the `+` seem to be doing in the code above?


## Topic 2: a little more about ***tibbles*** and ***types***

### (a) Without variable assignments, use the `%>%` "pipe" command to find the `class()` of `tibble::tibble(col=c(1,2,3))`

```{r, eval=FALSE, echo=FALSE, include=TRUE}
tibble(col=c(1,2,3)) %>% class()
```

### (b) Without variable assignments, use the `%>%` "pipe" command to find out the `class()` of `read_csv()` and `read.csv()`.

```{r, eval=FALSE, echo=FALSE, include=TRUE, message=FALSE}
read_csv("coffee_ratings.csv") %>% class() # read.csv("coffee_ratings.csv") %>% class()
```


## Topic 3: a little more about ***factors***

### (a) Confirm the observed column data types using the following stackoverflow conversation.

- https://stackoverflow.com/questions/21125222/determine-the-data-types-of-a-data-frames-columns

```{r, eval=FALSE, echo=FALSE, include=TRUE, message=FALSE}
sapply(coffee_ratings, class)
```

### (b) Change column data types in the `coffee_ratings` data and confirm the change using the code provided below based on the following stackoverflow conversation.

- https://stackoverflow.com/questions/27668266/dplyr-change-many-data-types

```{r, eval=FALSE, echo=FALSE, include=TRUE, message=FALSE}
coffee_ratings <- coffee_ratings %>% mutate(across(where(is.character), as.factor))
coffee_ratings %>% glimpse()
```


## Topic 4: A little more about ***factors*** and ***boolean vector masks***

### (a) Create the variable `coffee_ratings_nlevels` which lists the number of factor levels of each column using the provided code based on the following stackoverflow conversation and `nlevels` documentation.

- https://stackoverflow.com/questions/27676404/list-all-factor-levels-of-a-data-frame
- https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/nlevels

```{r, eval=FALSE, echo=FALSE, include=TRUE, message=FALSE}
coffee_ratings_nlevels <- sapply(coffee_ratings, nlevels)
coffee_ratings_nlevels
```

(b) Create a boolean mask vector from the `coffee_ratings_nlevels` variable above which is `TRUE` for each column with two or more but less than ten factor levels and `FALSE` otherwise using the provided code below.

```{r, eval=FALSE, echo=FALSE, include=TRUE, message=FALSE}
coffee_ratings_nlevels_boolean_mask <- (coffee_ratings_nlevels > 1) & (coffee_ratings_nlevels < 10)
```

(c) List the factor levels of each column with two or more factor levels using a vector boolean mask code provided below.

```{r, eval=FALSE, echo=FALSE, include=TRUE, message=FALSE}
sapply(coffee_ratings, levels)[coffee_ratings_nlevels_boolean_mask]
```


